---
title: "Final Code"
output: pdf_document
urlcolor: blue
---
\textbf{Authors}: Matthew Smith, Joseph Doraid

\textbf{Data Preparation}:

```{r, include=TRUE}
data <- read.csv('cleaned.csv', row.names='X')
data$Franch_Val <- as.numeric(gsub(",","",data$Franch_Val))
data$Franch_Val <- as.integer(data$Franch_Val)
head(data)
```

\textbf{EDA}:

Distribution of response and predictor variables:
```{r, include=TRUE}

ns <- colnames(subset(data, select=-c(Team, Year, Playoffs, CSF, CF, Finals, Championship, MPC1, MPC2, MPC3)))

for (i in 1:length(ns)){
  #png(paste('./Figures/', ns[i], '.png', sep=''))
  hist(data[ns[i]])
  #dev.off()
}
```

Outliers for operating income:
```{r, include=TRUE}
data[order(data$OI),]
```

Outliers for APG:
```{r, include=TRUE}
data[order(data$APG),]
```

Correlation Matrix:
```{r, include=TRUE}
round(cor(subset(data, select=-c(Team, Playoffs, CSF, CF, Finals, Championship, Year))), 2)
```


Plotted relationships between ***OI***, ***Ticket***, ***Followers***,  and ***Revenue***:
```{r, include=TRUE}
#png('./Figures/tickets_v_rev.png')
plot(data$Revenue ~ data$Ticket, main="Ticket Revenue vs. Total Revenue", xlab='Ticket Revenue (in million USD)', ylab='Total Revenue (in million USD)')
#dev.off()

#png('./Figures/oi_v_rev.png')
plot(data$Revenue ~ data$OI, main="Operating Income vs. Total Revenue", xlab='Operating Income (in million USD)', ylab='Total Revenue (in million USD)')
#dev.off()

#png('./Figures/followers_v_rev.png')
plot(data$Revenue ~ data$Followers, main="Social Media Followers vs. Total Revenue", xlab='Social Media Followers (in millions)', ylab='Total Revenue (in million USD)')
#dev.off()
```

Plotted relationships between ***Revenue***, ***Ticket***, ***Followers***, ***OI***, and ***Franch_Val***:
```{r, include=TRUE}
#png('./Figures/rev_v_fv.png')
plot(data$Franch_Val ~ data$Revenue, main="Total Revenue vs. Franchise Value", xlab='Total Revenue (in million USD)', ylab='Franchise Value (in million USD)')
#dev.off()

#png('./Figures/ticket_v_fv.png')
plot(data$Franch_Val ~ data$Ticket, main="Ticket Revenue vs. Franchise Value", xlab='Ticket Revenue (in million USD)', ylab='Franchise Value (in million USD)')
#dev.off()

#png('./Figures/oi_v_fv.png')
plot(data$Franch_Val ~ data$OI, main="Operating Income vs. Franchise Value", xlab='Operating Income (in million USD)', ylab='Franchise Value (in million USD)')
#dev.off()

#png('./Figures/followers_v_fv.png')
plot(data$Franch_Val ~ data$Followers, main="Social Media Followers vs. Franchise Value", xlab='Social Media Followers (in millions)', ylab='Franchise Value (in million USD)')
#dev.off()
```

```{r, include=TRUE}
teams <- unique(data$Team)
ymin <- min(data$Revenue) - 10
ymax <- max(data$Revenue) + 10
```

Plot total revenue vs time by team:
```{r, include=TRUE}
par(mar=c(5, 4, 4, 8), xpd=TRUE)

plot(data[data$Team == teams[1], ]$Revenue ~ data[data$Team == teams[1], ]$Year,
     type='l',
     ylim=c(ymin, ymax),
     col=1,
     xlab='Year',
     ylab='Total Revenue (in million USD)',
     main="Revenue vs. Time by Team")

for (i in 1:length(teams)){
  lines(data[data$Team == teams[i], ]$Revenue ~ data[data$Team == teams[i], ]$Year, lty=i, col=i)
}

legend("topright", inset=c(-0.3, 0), legend=teams, lty=1:length(teams), col=1:length(teams), pch=c(1,3), cex=0.5)
```

Plot total revenue vs time - Golden State Warriors:

```{r, include=TRUE}
plot(data[data$Team == 'Golden State Warriors', ]$Revenue ~ data[data$Team == 'Golden State Warriors', ]$Year,
     type='l',
     ylim=c(ymin, ymax),
     xlab='Year',
     ylab='Total Revenue (in million USD)',
     main="Revenue vs. Time - Golden State Warriors")

points(data[data$Team == 'Golden State Warriors', ]$Championship*data[data$Team == 'Golden State Warriors', ]$Revenue ~ data[data$Team == 'Golden State Warriors', ]$Year, col='red', pch=8)
```

Plot total revenue vs time - Miami Heat:

```{r, include=TRUE}
plot(data[data$Team == 'Miami Heat', ]$Revenue ~ data[data$Team == 'Miami Heat', ]$Year,
     type='l',
     ylim=c(ymin, ymax),
     xlab='Year',
     ylab='Total Revenue (in million USD)',
     main="Revenue vs. Time - Miami Heat")

points(data[data$Team == 'Miami Heat', ]$Championship*data[data$Team == 'Miami Heat', ]$Revenue ~ data[data$Team == 'Miami Heat', ]$Year, col='red', pch=8)
```

Plot total franchise value vs time by team:

```{r, include=TRUE}
teams <- unique(data$Team)
ymin_frv <- min(data$Franch_Val) + 5
ymax_frv <- max(data$Franch_Val) + 50
```

```{r, include=TRUE}
par(mar=c(5, 4, 4, 8), xpd=TRUE)

plot(data[data$Team == teams[1], ]$Franch_Val ~ data[data$Team == teams[1], ]$Year,
     type='l',
     ylim=c(ymin_frv, ymax_frv),
     col=1,
     xlab='Year',
     ylab='Franchise Value (in million USD)',
     main="Franchise Value vs. Time by Team")

for (i in 1:length(teams)){
  lines(data[data$Team == teams[i], ]$Franch_Val ~ data[data$Team == teams[i], ]$Year, lty=i, col=i)
}

legend("topright", inset=c(-0.3, 0), legend=teams, lty=1:length(teams), col=1:length(teams), pch=c(1,3), cex=0.5)
```

Plot total franchise value vs time - Golden State Warriors:

```{r, include=TRUE}
plot(data[data$Team == 'Golden State Warriors', ]$Franch_Val ~ data[data$Team == 'Golden State Warriors', ]$Year,
     type='l',
     ylim=c(ymin_frv+50, ymax_frv),
     xlab='Year',
     ylab='Franchise Value (in million USD)',
     main="Franchise Value vs. Time - Golden State Warriors")

points(data[data$Team == 'Golden State Warriors', ]$Championship*data[data$Team == 'Golden State Warriors', ]$Franch_Val ~ data[data$Team == 'Golden State Warriors', ]$Year, col='red', pch=8)
```

Plot total franchise value vs time - Miami Heat:

```{r, include=TRUE}
plot(data[data$Team == 'Miami Heat', ]$Franch_Val ~ data[data$Team == 'Miami Heat', ]$Year,
     type='l',
     ylim=c(ymin_frv+50, ymax_frv),
     xlab='Year',
     ylab='Franchise Value (in million USD)',
     main="Franchise Value vs. Time - Miami Heat")

points(data[data$Team == 'Miami Heat', ]$Championship*data[data$Team == 'Miami Heat', ]$Franch_Val ~ data[data$Team == 'Miami Heat', ]$Year, col='red', pch=8)
```

Significance testing - winning championships:

```{r, include=TRUE}
require(data.table)

n = 9

pvals <- rep(0, n)
samples <- rep(0, n)

results <- data.frame(Lag  = 1:n)

DT <- as.data.table(data)
```


```{r, include=TRUE}
for (i in 1:n) {
  DT$RL <- DT[, .(RL=shift(.(Revenue), type="lead", n=i)[[1L]]), by=Team]$RL
  DT$PC <- DT$RL / DT$Revenue
  
  png(paste('./Figures/rev_lag_', as.character(i), '.png', sep=''))
  boxplot(DT$PC ~ DT$Championship,
          main=paste('Future Revenue vs. Championship, Lag =', as.character(i)),
          xlab='Championship indicator in year t',
          ylab='Percentage change in revenue at year t + Lag')
  dev.off()
  
  pvals[i] <- t.test(DT$PC ~ DT$Championship, alternative='less')[3]$p.value
  samples[i] <- sum(!is.na(DT$PC))
}
```

```{r, include=TRUE}
results[['Number of Samples']] <- samples
results[['Number of Championship Samples']] <- samples/30
results[['P-Values, Revenue']] <- pvals
results[['Reject Null, Revenue']] <- pvals < 0.05 / n
```

```{r, include=TRUE}
for (i in 1:n) {
  DT$FVL <- DT[, .(FVL=shift(.(Franch_Val), type="lead", n=i)[[1L]]), by=Team]$FVL
  DT$PC_FVL <- DT$FVL / DT$Franch_Val
  
  png(paste('./Figures/fvl_lag_', as.character(i), '.png', sep=''))
  boxplot(DT$PC_FVL ~ DT$Championship,
          main=paste('Future Franchise Value vs Championship, Lag =', as.character(i)),
          xlab='Championship indicator in year t',
          ylab='Percentage change in franchise value at year t + Lag')
  dev.off()
  
  pvals[i] <- t.test(DT$PC_FVL ~ DT$Championship, alternative='less')[3]$p.value
}
```

```{r, include=TRUE}
results[['P-Values, Franchise Value']] <- pvals
results[['Reject Null, Franchise Value']] <- pvals < 0.05 / n
```

```{r, include=TRUE}
results
```

\textbf{Data Preparation - Part 2}:

```{r, include=TRUE}
data$FR <- DT[, .(RL=shift(.(Revenue), type="lead", n=1)[[1L]]), by=Team]$RL
data$PC <- data$FR / data$Revenue

data$FVL <- DT[, .(FVL=shift(.(Franch_Val), type="lead", n=1)[[1L]]), by=Team]$FVL
data$PC_FVL <- data$FVL / data$Franch_Val

data
```

Histogram of PC:
```{r, include=TRUE}
hist(data$PC)
```

Histogram of PC_FVL:
```{r, include=TRUE}
hist(data$PC_FVL)
```

Log-Transformations:
```{r, include=TRUE}
data$PC_FVL <- log(data$PC_FVL)
data$Revenue <- log(data$Revenue)
data$Ticket <- log(data$Ticket)
data$Population <- log(data$Population)
data$Income <- log(data$Income)
data$Followers <- log(data$Followers)
data$Franch_Val <- log(data$Franch_Val)
```

Set COVID indicator:
```{r, include=TRUE}
data$COVID <- as.numeric(data$Year == 2020)
data
```

Teams for Validation Set:
```{r, include=TRUE}
teams.val <- sample(teams, 3)
teams.val
```

Training Data:
```{r, include=TRUE}
data.train <- data[!(data$Team %in% teams.val) & (data$Year != 2021), ]
data.train.revenue <- subset(data.train, select = -c(FVL, PC_FVL))
row.names(data.train.revenue) <- NULL
tail(data.train.revenue)
```

```{r, include=TRUE}
data.train.franch.val <- subset(data.train, select = -c(FR, PC))
row.names(data.train.franch.val) <- NULL
```

```{r, include=TRUE}
data.val <- data[(data$Team %in% teams.val) & (data$Year != 2021), ]
data.val
```

\textbf{Baseline Models}:

```{r, include=TRUE}
baseline1 <- lm(PC ~ . - FR - Team, data=data.train.revenue)
```

```{r, include=TRUE}
plot(baseline1)
```
```{r, include=TRUE}
data.train.revenue[c(12, 260),]
```

```{r, include=TRUE}
data.train.revenue <- data.train.revenue[-c(12, 260),]
row.names(data.train.revenue) <- NULL
```

```{r, include=TRUE}
baseline1 <- lm(PC ~ . - FR - Team, data=data.train.revenue)

summary(baseline1)

print('BIC:')
print(BIC(baseline1))
```

```{r, include=TRUE}
baseline2 <- lm(PC_FVL ~ . - FVL - Team, data=data.train.franch.val)
```

```{r, include=TRUE}
plot(baseline2)
```

```{r, include=TRUE}
data.train.franch.val[c(222, 114),]
```

```{r, include=TRUE}
data.train.franch.val <- data.train.franch.val[-c(222, 114),]
row.names(data.train.franch.val) <- NULL
```

```{r, include=TRUE}
baseline2 <- lm(PC_FVL ~ . - FVL - Team, data=data.train.franch.val)

summary(baseline2)

print('BIC:')
print(BIC(baseline2))
```

\textbf{Intercept Models}:

```{r, include=TRUE}
intercept1 <- lm(PC ~ 1, data=data.train.revenue)
summary(intercept1)
print('BIC:')
print(BIC(intercept1))
```

```{r, include=TRUE}
intercept2 <- lm(PC_FVL ~ 1, data=data.train.franch.val)
summary(intercept2)
print('BIC:')
print(BIC(intercept2))
```

\textbf{Full Models}: 

```{r, include=TRUE}
full1 <- lm(PC ~ (.-FR-Team)^2, data=data.train.revenue)
summary(full1)
print('BIC:')
print(BIC(full1))
```
```{r, include=TRUE}
full2 <- lm(PC_FVL ~ (.-FVL-Team)^2, data=data.train.franch.val)
summary(full2)
print('BIC:')
print(BIC(full2))
```

\textbf{Stepwise Models}

```{r, include=TRUE}
stepwise1 = step(intercept1,
                 scope=list(lower = formula(intercept1), upper = formula(full1)),
                 direction = "both", trace = 0)
summary(stepwise1)
print('BIC:')
print(BIC(stepwise1))
```

```{r, include=TRUE}
stepwise2 = step(intercept2,
                 scope=list(lower = formula(intercept2), upper = formula(full2)),
                 direction = "both", trace = 0)
summary(stepwise2)
print('BIC:')
print(BIC(stepwise2))
```

\textbf{Mixed Effects Models}

```{r, include=TRUE}
require(lme4)
require(modelsummary)
```

```{r, include=TRUE}
mm0.pc <- lmer(PC ~ (1 | Year), data=data.train.revenue, REML=FALSE)
summary(mm0.pc)

print('BIC:')
print(BIC(mm0.pc))

print('Conditional R Squared:')
cor(predict(mm0.pc, data.train.revenue), data.train.revenue$PC)^2
```

```{r, include=TRUE}
mm0.fvl <- lmer(PC_FVL ~ (1 | Year), data=data.train.franch.val, REML=FALSE)
summary(mm0.fvl)

print('BIC:')
print(BIC(mm0.fvl))

print('Conditional R Squared:')
cor(predict(mm0.fvl, data.train.franch.val), data.train.franch.val$PC_FVL)^2
```

```{r, include=TRUE}
mm1.fvl <- lmer(PC_FVL ~ Income + WP + (1 | Year), data=data.train.franch.val, REML=FALSE)
summary(mm1.fvl)

print('BIC:')
print(BIC(mm1.fvl))

print('Conditional R Squared:')
cor(predict(mm1.fvl, data.train.franch.val), data.train.franch.val$PC_FVL)^2
```

\textbf{Validation}:

```{r, include=TRUE}
cor(predict(mm0.pc, data.val), data.val$PC)^2
```

```{r, include=TRUE}
cor(predict(mm1.fvl, data.val), data.val$PC_FVL)^2
```



